|

### 2024-10-28 output/seminar-slideshow.qmd
Make notation more consistent with article

$$
I_1 = \tau_1 + \lambda_{11} F_1^p + \lambda_{21} F_2^p + ... + \lambda_{n1} F_n^p + U_1^p \\
\vdots                                                                            \\
I_k = \tau_k + \lambda_{1k} F_1^p + \lambda_{2k} F_2^p + ... + \lambda_{nk} F_n^p + U_k^p
$$

@@

::: fragment
$$
I_1 = \tau_1 + \mathbf{\lambda}_{1}^T \mathbf{F} + U_1
$$
:::

@@

::::::::: columns
::::: {.column width="40%"}
$$
  i_k^* = \tau_k + \mathbf{\lambda}_k^T \mathbf{F} + U_k
$$

@@

::: {.fragment fragment-index="1"}
$$
  I_k = \left\{ \begin{matrix}
      1 &si &i_k^* \geq 0\\
      0 &si &i_k^* <    0
    \end{matrix} \right.
$$
:::

@@

::: {.fragment fragment-index="2"}
$$
  P(I_k = 1) = \int_{-\infty}^0 \mathcal{N}(i_k^*) \delta i_k^* \\
             = \Phi_N \left(\frac{\tau_k + \mathbf{\lambda}_k^T \mathbf{F}}{\sqrt{var(U_k)}}\right)
$$
:::
:::::

@@

# Plot: Upper Panel - Linear regression with Gaussian error distributions
latent_response |>
  ggplot(
    aes(x = response, y = point, fill = after_stat(x) > 0, group = point)
  )                            +
  geom_density_ridges_gradient(
    scale          =  .2,
    rel_min_height = 0.001,
    color          = PALETTE_BLUE["normal"]
  )                              +
  labs(x = "i~k~^*^", y = NULL)  +
  scale_y_continuous(
    breaks       = ERROR_POINTS,
    minor_breaks = NULL,
    labels       = NULL,
    limits       = range(ERROR_POINTS) * 1.12,
    expand       = expansion()
  )           +
  scale_x_continuous(breaks = 0) +
  scale_fill_manual(
    values = c(
      `FALSE` = unname(PALETTE_BLUE["white"]),
      `TRUE`  = unname(PALETTE_BLUE["normal"])
    )
  )                              +
  theme(
    axis.ticks.x = element_blank(),
    axis.title.y = element_markdown(angle = 90),
    legend.position = "none",
    panel.grid.major = element_line(linetype = "17")
  )                              +
  coord_flip()
```
:::

@@

# Plot: Lower Panel - Probit curve
probit_curve_data |>
  ggplot(mapping = aes(x = score, y = prob))                        +
  geom_line(color = PALETTE_BLUE["normal"], linewidth = LINE_WIDTH) +
  labs(x = "E(*i~k~^*^*)", y = "P(I~k~ = 1)")                       +
  scale_y_continuous(
    breaks       = 0:1,
    minor_breaks = NULL,
    limits       = 0:1,
    expand       = expansion(mult = c(0, .05))
  )  +
  scale_x_continuous(
    breaks       = ERROR_POINTS,
    minor_breaks = NULL,
    limits = range(ERROR_POINTS) * 1.12,
    expand = expansion()
  ) +
  theme(
    axis.title       = element_markdown(),
    axis.title.y     = element_markdown(angle = 90),
    panel.grid.major = element_line(linetype = "17"),
  )
```
:::
:::::
:::::::::

@@

logit_curve_data_long |>
  ggplot(
    mapping = aes(
      latent_trait, prob,
      color = item,
      group = after_scale(color)
    )
  )                                        +
  geom_line(linewidth = LINE_WIDTH)        +
  labs(x = "*&theta;*", y = "P(I~k~ = 1)") +
  scale_y_continuous(
    breaks       = 0:1,
    minor_breaks = NULL,
    limits       = 0:1,
    expand       = expansion(mult = c(0, .05))
  )                                        +
  scale_x_continuous(
    breaks       = ERROR_POINTS,
    minor_breaks = NULL,
    limits = range(ERROR_POINTS) * 1.12,
    expand = expansion()
  )                                        +
  scale_color_manual(values = ITEMS_PALETTE, guide = NULL) +
  theme(
    axis.title       = element_markdown(),
    axis.title.y     = element_markdown(angle = 90),
    panel.grid.major = element_line(linetype = "17"),
  )
```

@@

::::: {.column width="40%"}
::: {.fragment fragment-index="1"}
$$
  P(I_k = 1) = \Phi_L (a_k (\theta - b_k))
$$
:::

@@

::: {.fragment fragment-index="2"}
$$
  \begin{align}
    d_k &= -a_k b_k \\
    P(I_k = 1) &= \Phi_L (a_k \theta + d_k))
  \end{align}
$$
:::

@@

$$
  P(I_k = 1) = \Phi_L (\mathbf{a}_k^T \mathbf{\theta} + d_k)
$$

@@

plot_ly(
  x = ~PROB_AXIS_2D,
  y = ~PROB_AXIS_2D,
  z = ~mirt_prob_surface,
  colorscale = list(c(0, 1), PALETTE_BLUE["normal"] |> rep(2)),
  type = 'surface',
  hovertemplate = 'P(I<sub>k</sub> = 1) = %{z:.2f}<extra></extra>',
  showscale = FALSE,
  hoverinfo = 'Z'
) |>
  layout(
    scene = list(
      xaxis = list(title = "<i>θ</i><sub>1</sub>"),
      yaxis = list(title = "<i>θ</i><sub>2</sub>"),
      zaxis = list(title = "P(I<sub>k</sub> = 1)"),
      camera = list(
        center = list(x = -.1, z = -.2),
        eye = list(x = -.12, y = -1.8, z = 1)
      )
    )
  ) |>
  config(displayModeBar = FALSE, scrollZoom = FALSE)
```
:::

@@

::: columns
::: column
$$
  P(I_k = 1) = \Phi_L (\mathbf{a}_k^T \mathbf{\theta} + d_k)
$$

@@

::: fragment
$$
  \begin{align}
    P(I_k = 1) &= \Phi_L \left([1, 0.5]
      \begin{bmatrix}
         1 \\
        -1
      \end{bmatrix}
      -0.5 \right) \\
      &= \Phi_L \left(1 · 1 + 0.5 · (-1) -0.5 \right) \\
      &= \Phi_L \left(1 -0.5 -0.5 \right) = \Phi_L \left(0 \right) \\
      &= \frac{1}{1 + e^{-0}} = \frac{1}{1 + 1} = \frac{1}{2} = 0.5
  \end{align}
$$
:::
:::
:::

@@

$$
  \begin{align}
    P(I_k = 1) &= \Phi_N \left(
      \frac{\mathbf{\lambda}_k^T \mathbf{F} + \tau_k}{\sqrt{var(U_k)}}
    \right) \\
    P(I_k = 1) &= \Phi_L (\mathbf{a}_k^T \mathbf{\theta} + d_k)
  \end{align}
$$

@@

$$
  \begin{align}
    \mathbf{a}_k &= \frac{1.7}{\sqrt{var(U_k)}} \mathbf{\lambda}_k \\
    d_k &= \frac{1.7}{\sqrt{var(U_k)}} \tau_k
  \end{align}
$$
:::

@@

$$
  P(I_k = 1) = \Phi_L (\mathbf{a}_k^T \mathbf{\theta} + d_k)
$$

::: fragment
$$
  P(I_k = 1) = \Phi_L (a_k (\theta - b_k))
$$
:::

@@

$$
MID = \frac{-d_k}{\sqrt{\mathbf{a}_k^T\mathbf{a}_k}}
$$

@@

```{r mid-representation}
#| fig-height: 8
#| fig-width:  8

@@

contour_plot +
    geom_segment(
    data    = mirt_item_spikes,
    mapping = aes(
      trait_1_start, trait_2_start,
      xend = trait_1_end,
      yend = trait_2_end
    ),
    inherit.aes = FALSE,
    linewidth = LINE_WIDTH,
    linetype  = '43',
    color = PALETTE_BLUE["light"]
  ) +
  geom_point(
    data = mirt_item_coords,
    mapping = aes(origin_transf_1, origin_transf_2),
    inherit.aes = FALSE,
    size = I(4)
  ) +
  scale_x_continuous(
    breaks = -1:1,
    minor_breaks = NULL,
    limits = c(-1.5, 1.5)
  ) +
  scale_y_continuous(
    breaks = -1:1,
    minor_breaks = NULL,
    limits = c(-1.5, 1.5)
  ) +
  annotate(
    geom = "text",
    x = 0.1, y = 0.2,
    label = "MID",
    size = 6,
    angle = mirt_item_params |> pull(deg_1)
  )
```

<!--# TODO: Fórmula de los cosenos directores -->

>>>>

$$
Y_1 = \tau_1 + \lambda_{11} F_1^p + \lambda_{21} F_2^p + ... + \lambda_{n1} F_n^p + U_1^p \\
\vdots                                                                            \\
Y_i = \tau_i + \lambda_{1i} F_1^p + \lambda_{2i} F_2^p + ... + \lambda_{ni} F_n^p + U_i^p
$$

@@

::: fragment
$$
Y_i = \tau_i + \mathbf{\lambda}_i^T \mathbf{F} + U_i
$$
:::

@@

::::::::: columns
::::: {.column width="40%"}
$$
  y_i^* = \tau_i + \mathbf{\lambda}_i^T \mathbf{F} + U_i
$$

@@

::: {.fragment fragment-index="1"}
$$
  Y_i = \left\{ \begin{matrix}
      1 &si &y_i^* \geq 0\\
      0 &si &y_i^* <    0
    \end{matrix} \right.
$$
:::

@@

::: {.fragment fragment-index="2"}
$$
  P(Y_i = 1) = \int_{-\infty}^0 \mathcal{N}(y_i^*) \delta y_i^* \\
             = \Phi_N \left(\frac{\tau_i + \mathbf{\lambda}_i^T \mathbf{F}}{\sqrt{var(U_i)}}\right)
$$
:::
:::::

@@

# Plot: Upper Panel - Linear regression with Gaussian error distributions
latent_response |>
  ggplot(
    aes(x = response, y = point, fill = after_stat(x) > 0, group = point)
  )                            +
  geom_density_ridges_gradient(
    scale          =  .2,
    rel_min_height = 0.001,
    color          = PALETTE_BLUE["normal"]
  )                              +
  labs(x = "y~i~^*^", y = NULL)  +
  scale_y_continuous(
    breaks       = ERROR_POINTS,
    minor_breaks = NULL,
    labels       = NULL,
    limits       = range(ERROR_POINTS) * 1.12,
    expand       = expansion()
  )           +
  scale_x_continuous(breaks = 0) +
  scale_fill_manual(
    values = c(
      `FALSE` = unname(PALETTE_BLUE["white"]),
      `TRUE`  = unname(PALETTE_BLUE["normal"])
    )
  )                              +
  theme(
    axis.ticks.x = element_blank(),
    axis.title.y = element_markdown(angle = 90),
    legend.position = "none",
    panel.grid.major = element_line(linetype = "17")
  )                              +
  coord_flip()
```
:::

@@

# Plot: Lower Panel - Probit curve
probit_curve_data |>
  ggplot(mapping = aes(x = score, y = prob))                        +
  geom_line(color = PALETTE_BLUE["normal"], linewidth = LINE_WIDTH) +
  labs(x = "E(*y~i~^*^*)", y = "P(Y~i~ = 1)")                       +
  scale_y_continuous(
    breaks       = 0:1,
    minor_breaks = NULL,
    limits       = 0:1,
    expand       = expansion(mult = c(0, .05))
  )  +
  scale_x_continuous(
    breaks       = ERROR_POINTS,
    minor_breaks = NULL,
    limits = range(ERROR_POINTS) * 1.12,
    expand = expansion()
  ) +
  theme(
    axis.title       = element_markdown(),
    axis.title.y     = element_markdown(angle = 90),
    panel.grid.major = element_line(linetype = "17"),
  )
```
:::
:::::
:::::::::

@@

logit_curve_data_long |>
  ggplot(
    mapping = aes(
      latent_trait, prob,
      color = item,
      group = after_scale(color)
    )
  )                                        +
  geom_line(linewidth = LINE_WIDTH)        +
  labs(x = "*&theta;*", y = "P(Y~i~ = 1)") +
  scale_y_continuous(
    breaks       = 0:1,
    minor_breaks = NULL,
    limits       = 0:1,
    expand       = expansion(mult = c(0, .05))
  )                                        +
  scale_x_continuous(
    breaks       = ERROR_POINTS,
    minor_breaks = NULL,
    limits = range(ERROR_POINTS) * 1.12,
    expand = expansion()
  )                                        +
  scale_color_manual(values = ITEMS_PALETTE, guide = NULL) +
  theme(
    axis.title       = element_markdown(),
    axis.title.y     = element_markdown(angle = 90),
    panel.grid.major = element_line(linetype = "17"),
  )
```

@@

::::: {.column width="40%"}
::: {.fragment fragment-index="1"}
$$
  P(Y_i = 1) = \Phi_L (a_i (\theta - b_i))
$$
:::

@@

::: {.fragment fragment-index="2"}
$$
  \begin{align}
    d_i &= -a_i b_i \\
    P(Y_i = 1) &= \Phi_L (a_i \theta + d_i))
  \end{align}
$$
:::

@@

$$
  P(Y_i = 1) = \Phi_L (\mathbf{a}_i^T \mathbf{\theta} + d_i)
$$

@@

plot_ly(
  x = ~PROB_AXIS_2D,
  y = ~PROB_AXIS_2D,
  z = ~mirt_prob_surface,
  colorscale = list(c(0, 1), PALETTE_BLUE["normal"] |> rep(2)),
  type = 'surface',
  hovertemplate = 'P(I<sub>i</sub> = 1) = %{z:.2f}<extra></extra>',
  showscale = FALSE,
  hoverinfo = 'Z'
) |>
  layout(
    scene = list(
      xaxis = list(title = "<i>θ</i><sub>1</sub>"),
      yaxis = list(title = "<i>θ</i><sub>2</sub>"),
      zaxis = list(title = "P(I<sub>i</sub> = 1)"),
      camera = list(
        center = list(x = -.1, z = -.2),
        eye = list(x = -.12, y = -1.8, z = 1)
      )
    )
  ) |>
  config(displayModeBar = FALSE, scrollZoom = FALSE)
```
:::

@@

::: columns
::: column
$$
  P(Y_1 = 1) = \Phi_L (\mathbf{a}_1^T \mathbf{\theta} + d_1)
$$

@@

::: fragment
$$
  \begin{align}
    P(Y_1 = 1) &= \Phi_L \left([1, 0.5]
      \begin{bmatrix}
         1 \\
        -1
      \end{bmatrix}
      -0.5 \right) \\
      &= \Phi_L \left(1 · 1 + 0.5 · (-1) -0.5 \right) \\
      &= \Phi_L \left(1 -0.5 -0.5 \right) = \Phi_L \left(0 \right) \\
      &= \frac{1}{1 + e^{-0}} = \frac{1}{1 + 1} = \frac{1}{2} = 0.5
  \end{align}
$$
:::
:::
:::

@@

$$
  \begin{align}
    P(Y_i = 1) &= \Phi_N \left(
      \frac{\mathbf{\lambda}_i^T \mathbf{F} + \tau_i}{\sqrt{var(U_i)}}
    \right) \\
    P(Y_i = 1) &= \Phi_L (\mathbf{a}_i^T \mathbf{\theta} + d_i)
  \end{align}
$$

@@

$$
  \begin{align}
    \mathbf{a}_i &= \frac{1.7}{\sqrt{var(U_i)}} \mathbf{\lambda}_i \\
    d_i &= \frac{1.7}{\sqrt{var(U_i)}} \tau_i
  \end{align}
$$
:::

@@

$$
  P(Y_i = 1) = \Phi_L (\mathbf{a}_i^T \mathbf{\theta} + d_i)
$$

::: fragment
$$
  P(Y_i = 1) = \Phi_L (a_i (\theta - b_i))
$$
:::

@@

$$
  MID = \frac{-d_i}{\sqrt{\mathbf{a}_i^T\mathbf{a}_i}}
$$

@@

:::: columns
::: column
```{r mid-representation}
#| fig-height: 8
#| fig-width:  8

@@

contour_plot +
    geom_segment(
    data    = mirt_item_spikes,
    mapping = aes(
      trait_1_start, trait_2_start,
      xend = trait_1_end,
      yend = trait_2_end
    ),
    inherit.aes = FALSE,
    linewidth = LINE_WIDTH,
    linetype  = '43',
    color = PALETTE_BLUE["light"]
  ) +
  geom_point(
    data = mirt_item_coords,
    mapping = aes(origin_transf_1, origin_transf_2),
    inherit.aes = FALSE,
    size = I(4)
  ) +
  scale_x_continuous(
    breaks = -1:1,
    minor_breaks = NULL,
    limits = c(-1.5, 1.5)
  ) +
  scale_y_continuous(
    breaks = -1:1,
    minor_breaks = NULL,
    limits = c(-1.5, 1.5)
  ) +
  annotate(
    geom = "text",
    x = 0.1, y = 0.2,
    label = "MID",
    size = 6,
    angle = mirt_item_params |> pull(deg_1)
  )
```
:::

::: column
::: fragment
$$
  cos \, \mathbf{\alpha}_i = \frac{\mathbf{a}_i}{\sqrt{\mathbf{a}_i^T\mathbf{a}_i}}
$$
:::

::: fragment

:::
:::
::::
